---
title: "About Page"
author: "Leena Ambady"
date: "3/7/2020"
output: html_document
---
Url to Repository: https://github.com/lambady/Final-Project-2-25-20.git 

For my final project, I want to look at rates of organ donation registration in New York. More specifically, I want to see whether organ procurement organizations or demographic factors have an affect on rates of registration. Organ procurement organizations are supposed to encourage registration and donation and education the counties that they serve. 

This week, I was able to make progress on getting and cleaning some of the demographic data I'm interested in using. I got data from the 2015 American Communities Survey. I was able to get county level racial/ethnic data for New York. I then joined this data with an existing dataset I had that gave organ donation registration enrollments for each county in New York. This data was collected on a monthly basis from the years 2008-2015. Obviously, there are quite a lot of observations then. Something I need to figure out is how to reconcile this registration data was collected very frequently with he demographic data that was collected once every few years. For this week, I just focused on joining the data for the year 2015, since I had ACS data for that. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
dir.create("raw-data")
library(readr)
library(tidyverse)
library(janitor)
library(dplyr)
library(readxl)
library(tidycensus)
library(stringr)
census_api_key("41e31e7dc9f4be0cac86dd847ae1e2e19837090b")
library(broom)
library(forcats)
```

```{r Reading in data}
# This dataset contains the number of registry enrollments in each county of New
# York for the years 2008-2015

new_york <- read.csv("raw-data/Donate_Life_Organ_and_Tissue_Donor_Registry_Enrollment_by_County__Beginning_September_2008.csv") %>%
  clean_names()

```

```{r Cleaning New York Dataset}
new_york_cleaned <- rename(new_york, 
                           "2012 Census Population" = x2012_census_population,
                           "% Eligible Population Enrolled" = x_eligible_population_enrolled, 
                           "Population 18+ Estimate" = population_18_estimate, 
                           "Organ Procurement Organization" = opo, 
                           "Year" = year, "Month" = month, 
                           "County" = county, 
                           "Registry Enrollments" = registry_enrollments)
```

```{r Getting ACS Information that is of interest, include = FALSE}
#variables <- load_variables(2015, "acs5")

# This code gives me census data from 2015 about the differe ethnic/racial
# demographics in all the different counties of New York. I hope to join this
# with the county data that I have collected on rates of organ donor
# registration in each county of New York.

# racevars <- c(White = "B02001_002", 
#               Black = "B02001_003", 
#               Asian = "B02001_005",
#               Hispanic = "B03003_003", 
#               Median_income = "B06011_001",
#               Median_age = "B01002_001") 

ny_county_data <- get_acs(geography = "county",
                  variables =  c(White = "B02001_002",
              Black = "B02001_003",
              Asian = "B02001_005",
              Hispanic = "B03003_003"),
                  year = 2015,
                  state = "NY",
                  geometry = TRUE,
                  summary_var = "B02001_001")

ny_income <- get_acs(geography = "county",
                  variables = c(Median_income = "B19013_001"),
                  year = 2015,
                  state = "NY",
                  geometry = TRUE)

ny_age <- get_acs(geography = "county",
                  variables = c(Median_age = "B01002_001"),
                  year = 2015,
                  state = "NY",
                  geometry = TRUE)


```

```{r}
# One thing I want to do is link the census data for each county to the data
# about the number of registry enrollments in each county. I can do this by
# joining the new_york_cleaned and ny_county_data datasets. First, I need to
# alter the ny_county_data so that it can be joined. Currently the County names
# in this dataset are formatting like "Allegheny County, New York" whereas in
# the new_york_cleaned they are just "Allegheny." Below, I created a new column
# with just the first word so that I can join the datasets.

ny_county_data$County <- word(ny_county_data$NAME, 1) 
ny_county_data_final <- ny_county_data %>%
  mutate(County = recode(County, New = "New York", "St." = "St Lawrence")) 

# Another thing is that the ny_county_data demographic is just from 2015,
# whereas the registry data was recorded each month from 2008-2015. Right now, I
# will just filter it for 2015, but eventually I will need to figure out what
# dates I want to focus on, and if I want to get more demographic data from
# other ACS years. I am not sure it makes sense to have monthly data, and might
# just focus on different years.

County_enrollment_demographic_data <- new_york_cleaned %>%
  filter(Year == "2015") %>%
  filter(!County == "Out of State") %>%
  filter(!County == "Unknown") %>%
  filter(!County == "TOTAL NYS") %>%
  mutate(County = recode(County, Cattauragus = "Cattaraugus")) %>%
  full_join(ny_county_data_final, by = "County") 

# Now I have one dataset with information on the number of enrollment registrations and the demographic data for each county in New York. 
```

```{r}
Simplified_Joined <- County_enrollment_demographic_data %>%
  select(-"Organ Procurement Organization", 
         -"location", -"chart_month", -"GEOID", 
         -"NAME", -"geometry") %>%
  rename("Pop_Enrolled" = "% Eligible Population Enrolled") %>%
  mutate(prop = estimate/summary_est)
  


Simplified_Joined %>%
  filter(Month == "11") %>%
  filter(!variable == "White") %>%
  ggplot(aes(x = Pop_Enrolled, 
             y = fct_reorder(County, Pop_Enrolled), 
             color = prop)) + 
  facet_wrap(~ variable) + 
  geom_point(alpha = 0.7) +
  theme(axis.text.y = element_text(size = 5.5)) +
    scale_colour_gradient(low = "yellow", high = "red") + 
  labs(x = "% of Population Registered as an Organ Donor", 
       y = "County",
       title = "New York Organ Donation Registration Rates by County in 2015",
       subtitle = "Registration Rates Measured Monthly",
       color = "Percent") 

ggsave("MS6/prop_white.png")
  
```


```{r}
Simplified_Joined %>%
  filter(Month == "11") %>%
  filter(variable == "Median_age") %>%
  ggplot(aes(x = estimate, y = Pop_Enrolled, label = County)) +
  geom_point() +
  geom_text(vjust = 0.5, nudge_y = 1, size = 2)

# could i get it so when you hover over the point the County name shows up? 
```

```{r}
Simplified_Joined %>%
   filter(Month == "11") %>%
  filter(variable == "Median_income") %>%
  ggplot(aes(x = estimate, y = Pop_Enrolled, label = County)) +
  geom_point() +
  geom_text(vjust = 0.5, nudge_y = 1, size = 2) 

Simplified_Joined %>%
filter(variable == "Median_income") %>%
filter(!is.na(estimate)) %>%
  summarize(correlation_income = cor(estimate, Pop_Enrolled))

Simplified_Joined %>%
filter(variable == "Median_age") %>%
filter(!is.na(estimate)) %>%
  summarize(correlation_income = cor(estimate, Pop_Enrolled))
  
```

```{r Registration Rates}
County_enrollment_demographic_data %>%
  rename("Pop_Enrolled" = "% Eligible Population Enrolled") %>%
  filter(Month == "11") %>%
  ggplot(aes(fill = Pop_Enrolled, geometry = geometry)) +
  geom_sf()

# want to get the county names to show up here
# I want to figure out a way to make similar maps but for income and age and in some way race...
```

```{r Income}
ny_income$County <- word(ny_income$NAME, 1) 
ny_income_final <- ny_income %>%
  mutate(County = recode(County, New = "New York", "St." = "St Lawrence"))
 
ny_income_final %>%
  ggplot(aes(fill = estimate, geometry = geometry)) +
  geom_sf()
  
  
```

```{r Age}
ny_age$County <- word(ny_income$NAME, 1) 
ny_age_final <- ny_age %>%
  mutate(County = recode(County, New = "New York", "St." = "St Lawrence"))
 
ny_age_final %>%
  ggplot(aes(fill = estimate, geometry = geometry)) +
  geom_sf()
  
```

```{r OPOs}
County_enrollment_demographic_data %>%
  rename("OPO" = "Organ Procurement Organization") %>%
  filter(Month == "11") %>%
  ggplot(aes(fill = OPO, geometry = geometry)) +
  geom_sf()

```

```{r Increase in Rates by OPO}
new_york_cleaned %>%
  filter(!County == "Out of State") %>%
  filter(!County == "Unknown") %>%
  filter(!County == "TOTAL NYS") %>%
    rename("Pop_Enrolled" = "% Eligible Population Enrolled") %>%
  rename("OPO" = "Organ Procurement Organization") %>%
  group_by(Year, OPO) %>%
  summarize(avg_registrants = mean(Pop_Enrolled)) %>%
  ggplot(aes(x = Year, y = avg_registrants, color = OPO)) + geom_line() 

```


```{r}
new_york_cleaned %>%
  filter(Month == "1") %>%
  filter(!County == "Out of State") %>%
  filter(!County == "Unknown") %>%
  filter(!County == "TOTAL NYS") %>%
  rename("Pop_Enrolled" = "% Eligible Population Enrolled") %>%
  pivot_wider(names_from = Year, values_from = Pop_Enrolled) %>%
  View()

# how do i consolidate this ??
```

How might I get the a map with the change/increase in registrants from 2007 to 2020
  probably need to do a pivot_wider to make each year a column and then subtract 

```{r Race Map}
County_enrollment_demographic_data %>%
  mutate(prop = estimate / summary_est) %>%
  ggplot(aes(fill = prop, color = prop, geometry = geometry)) +
  facet_wrap(~ variable) +
  geom_sf() + 
  scale_fill_viridis_c(direction = -1) +
  scale_color_viridis_c(direction = -1)

```

Could do increase faceted by race 

```{r}
# Was trying to do the increase in registration over time by race, but don't have race data for all counties. 
# Instead, want to try the increase over time but just make a category for each whether they are majority minority or not 


test <- new_york_cleaned %>%
  filter(!County == "Out of State") %>%
  filter(!County == "Unknown") %>%
  filter(!County == "TOTAL NYS") %>%
  mutate(County = recode(County, Cattauragus = "Cattaraugus")) %>%
  full_join(ny_county_data_final, by = "County") %>%
  mutate(percent = (estimate / summary_est) * 100) %>%
  filter(variable == "White") %>%
  mutate(binary = ifelse(percent <= 80, "minority", "white")) %>%
  rename("Pop_Enrolled" = "% Eligible Population Enrolled")

test_all <- new_york_cleaned %>%
  filter(!County == "Out of State") %>%
  filter(!County == "Unknown") %>%
  filter(!County == "TOTAL NYS") %>%
  mutate(County = recode(County, Cattauragus = "Cattaraugus")) %>%
  full_join(ny_county_data_final, by = "County") %>%
  mutate(prop = estimate / summary_est) %>%
  mutate(binary = ifelse(prop <= 0.8, "minority", "white")) %>%
  rename("Pop_Enrolled" = "% Eligible Population Enrolled") 


```


```{r}
# test %>%
#   filter(Month == "11") %>%
#   group_by(Year, binary) %>%
#   summarize(avg_registrants = mean(Pop_Enrolled)) %>%
#   ggplot(aes(x = Year, y = avg_registrants, color = binary)) + 
#   geom_line() 


test %>%
  group_by(binary) %>%
  summarize(mean = mean(Pop_Enrolled))

 test %>%
 ggplot(aes(x = Pop_Enrolled)) +
 geom_histogram() +
  facet_wrap(~ binary)


 model <- lm(Pop_Enrolled ~ binary, data = test)
 
 model %>%
   tidy(conf.int = TRUE) %>%
   select(term, estimate, conf.high, conf.low)
 
 model2 <- lm(Pop_Enrolled ~ percent, data = test)
 
 model2 %>%
    tidy(conf.int = TRUE) %>%
   select(term, estimate, conf.high, conf.low)
   
 test %>%
   filter(Month == "1") %>%
   filter(Year == "2015") %>%
   ggplot(aes(x = percent, y = Pop_Enrolled)) + geom_point() +
   geom_smooth(method = "lm", se = FALSE)
 
  # is this too skewed by the outliers ?? 
  
```
Then organize everything so it is not so messy 

Need to decide what graphs I want, what maps I want, and display them 

Need to figure out trust
